How has the Brazilian Amazon been constructed as a problem? An analysis of

### 2022-04-08 Draft_April_2022/Full_draft_20220401.rmd
Updated some statistics to be consistent with new findings and added some of the footnotes and references to text

Problem-construction at the level of discourse is also more varied. They are not monolithic in time, across location, or even by the same speaker. While governmental discourses in Brazil have been studied for topic such as inflation or race relations, we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

@@

We build upon the dataset provided by Cezar (2020) which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil since 1985. We then identify all speeches about the Amazon as a region, peoples, or forest out of the 6130 speeches. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

@@

We chose to employ a support-vector machine (SVM) algorithm to label texts, that is, a non-probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories (see Meyer et al. 2021; Noble 2006). After the SVM model is trained and tuned, and the model shows to appropriately fit the validation set, we use the model to automatically code the remaining 1007 Amazonian statements. The model was also trained to identify false positive observations, which we chose to delete for analysis since these are not meaningful statements about the Amazon. The final dataset for analysis contains 1895 coded Amazonian statements. Finally, we use poldis once more to extract locations for all speeches in the data. These locations represent the Brazilian state in which certain speech was given or an international country.

@@

To analyze our data, we first present a series of different plots on proportions of Amazonian statements and problem-constructions over time and by presidents. We control proportions for factors that might affect the incidence of the Amazon in speeches and portray both real and controlled curves. To test whether different problem constructions change according to location, we run a multinomial regression model in which different problem constructions (as categories) are the dependent variable and location is the independent variable. We also control for annual deforestation rates, annual inflation rates, and election years in the model. We interpret the plots and model considering multiple Amazon related events and policies over the last 30 years, as well as their correlations with different presidents and locations, while embedding problem-construction in contemporary happenings.

@@

[^2]: The reasoning here is that knowledge of high or low deforestation rates might drive presidents to speak about the Amazon. Matching deforestation with speech dates is a complicated endeavor. INPE published consolidated deforestation data with almost two years of delay until 2005 [@capobianco2021, pp. 60]. Preliminary data was published earlier, but with lower degrees of confidence in the findings. For the subsequent years, the consolidated figures for a given year tend to come out in August of the subsequent year. However, it also seems to be the case that the executive government have access to the data before everyone else. In addition, other sources indicating if deforestation is going up or down, as fire data or lower resolution deforestation data (DETER for example), circulates within the same year. The year of 1988 is particularly indicative of these complications: a report dating 1988 was circulating with deforestation and fire figures[@fearnside1990], and a New York Times article about the issue was written with comments by an INPE scientist [@simons1988]. Hence, we believe the most appropriate way of matching is without any lags on deforestation.

@@

We now move to mixed types, which average at around 17% for all presidents in our sample: overall, presidents prefer pure problem-constructions. While there is some variation in time for each single mixed type, some of them have low counts and interpretations are not adequate. We focus our discussion on those with higher incidence. First, the most frequent mix overall is that of economic integration with conservation, which after reaching 25% of all problem-constructions in 1989, remained stable at 9% on average for the remainder of the period. President Lula was the most frequent user of this mix. Second, except for economic and social development appearing together in all statements made by Color in 1992, mixed types using conservation were quite frequent in the lead up and aftermath of the 1992 Earth Summit. This includes the mix type we label sustainable development, which constructs the Amazon as a problem of economic integration, social development, and environmental conservation. In all, we interpret the appearance of mixed types as more complex understandings of Amazonian problems. This follows a global agenda of understanding interconnections of social, environmental, and economic domains. As we show that Amazonian incidence in discourse does respond to global issues, this is not a surprise given agendas as Millennium Development Goals and the Sustainable Development Goals. Nevertheless, as in pure types, we also observe the comeback of sovereignty being used in mixed types in detriment of conservation. This becomes more apparent in a comparison between Lula and Bolsonaro, the two presidents that mix the most with proportions 11% above presidential averages: 28.5% and 28.6% respectively. While the former frequently mixed conservation with other problem constructions, the latter prefers mixing sovereignty. The combination of sovereignty with economic integration, which was also characteristic of the military dictatorship policies for the region, reaches its highest level with Bolsonaro.

```{r Figure 4: mixed-types by president , echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mixed_types_presid <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(president, mx_cat) %>%
  group_by(president) %>%
  mutate(perc = n / sum(n)) %>%
  mutate(president = factor(president,
                            levels = c("Sarney", "Collor", "Itamar", "FHC", "Lula", "Dilma", "Temer", "Bolsonaro"))) %>%
  ggplot(aes(x=president, y=perc, fill=mx_cat,label= scales::percent(perc, accuracy=.1))) +
  geom_bar(stat="identity", position="stack") +
  geom_text(size = 3, position = position_stack(vjust=.5)) +
  scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "", title ="Share of each mixed type problem-constructions by president",
       subtitle = "Sample composed of Amazonian statements that were coded as a mixed-type (n=431)",
       caption = "") +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
mixed_types_presid
```

@@

```{r multimonial model, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
# data on deforestation rates per year (PRODES)
# source: http://terrabrasilis.dpi.inpe.br/app/dashboard/deforestation/biomes/legal_amazon/rates
# Since annual deforestation rates are calculated since 1988, we copy the rates of 1988 to 1985, 1986, and 1987.
# These rates are consistent with most estimates (see for example Prates and Bracha 2010 -
# https://www.scielo.br/j/ecos/a/59DMy3zmJdPHXzXRsHTkmNF/?lang=pt&format=pdf)
adr <- data.frame(year = c(1985, 1986, 1987), def_year = c(21.050, 21.050, 21.050))
amazon_def_year <- rbind(adr, amazon_def_year)
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
# inflation rates per year Brazil (average for year)
# source EU inflation: https://www.inflation.eu/en/inflation-rates/brazil/historic-inflation/cpi-inflation-brazil.aspx
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
# remove non-identified locations and mixed-other constructions (very little data for either and not relevant)
ama_model <- filter(ama_model, location_cat != "Non Identified", mx_cat != "Other Mixed-types") %>% 
  mutate(mx_cat = factor(mx_cat, levels = c("Pure Conservation", "Pure Economic Integration",
                                            "Pure Social Development", "Pure Sovereignty", "Economic Conservation",
                                            "Economic and Social Development", "Sovereign Conservation",
                                            "Sustainable Development","Sovereign Integration",
                                            "Social Conservation", "Other")),
         election_year = ifelse(grepl("1989|1994|1998|2002|2006|2010|2014|2018", year), 1, 0))
# model
model <- multinom(mx_cat ~ location_cat + election_year + def_year + AAI, data = ama_model, model = TRUE)
# summary statistics
model_summary <- summary(model)
p_values <- round((1 - pnorm(abs(summary(model)$coefficients/summary(model)$standard.errors)))*2, 5)
# post estimation tests
chi_square <- chisq.test(ama_model$mx_cat, predict(model)) 
# Chi-square p value (p-value < 0.001) tells us that our model explains a
# significant amount of the variation in the data in comparison with a null model.
# null_model <- multinom(mx_cat ~ 1, data = ama_model) # Null model for a likelihood ration test, just in case.
# library(lmtest)
# lrtest(null_model, model) # We very much reject the hypothesis
# (i.e. the more complex model is significantly better than the null model).
# Finally, let's get a nice output:
stargazer::stargazer(model, type="text",
                     covariate.labels = c("Brasilia", "International", "Non-AM state",
                                          "Election", "Deforestation", "Inflation"),
                     column.labels = c("EI", "SD", "Sov", "EI-Con", "EI-SD", "Sov-Con", "SD-EI-Con",
                                       "Sov-EI", "SD-Con", "Other"),
                     dep.var.labels.include = FALSE)
```

>>>>

Problem-construction at the level of discourse is also more varied. They are not monolithic in time, across location, or even by the same speaker. While governmental discourses in Brazil have been studied for topic such as inflation or race relations^[For an overview on how inflation and political discourse in Brazil see Porto (2007), Malheiros-Poulet (2012), and Fonseca et al. (2021). For a few studies on racist and anti-racist discourses in Brazil see Htun (2004), Silva and Larkins (2019) and Dijk (2020).], we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

@@

We build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil since 1985. We then identify all speeches about the Amazon as a region, peoples, or forest out of the 6130 speeches. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

@@

We chose to employ a support-vector machine (SVM) algorithm to label texts, that is, a non-probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories [@meyer2021; @noble2006]^[For more information on this please see the extended methodology. The document includes a detailed discussion of how the SVM algorithm was chosen over other algorithms based on evaluation of performance of trained models for labeling the  validation set. The extended methodology also includes a more detailed codebook table with further coding guidelines.]. After the SVM model is trained and tuned, and the model shows to appropriately fit the validation set, we use the model to automatically code the remaining 1007 Amazonian statements. The model was also trained to identify false positive observations, which we chose to delete for analysis since these are not meaningful statements about the Amazon. The final dataset for analysis contains 1895 coded Amazonian statements. Finally, we use poldis once more to extract locations for all speeches in the data. These locations represent the Brazilian state in which certain speech was given or an international country.

@@

To analyze our data, we first present a series of different plots on proportions of Amazonian statements and problem-constructions over time and by presidents. We control proportions for factors that might affect the incidence of the Amazon in speeches and portray both real and controlled curves. To test whether different problem constructions change according to location, we run a multinomial regression model in which different problem constructions (as categories) are the dependent variable and location is the independent variable. We code 4 different location categories for where speeches take place, they are: Amazonian states, non-Amazonian states, Brasilia, and International. In the model, we also control for annual deforestation rates, annual inflation rates, and election years in the model as these are variables which could mitigate the correlations between problem construction and location^[See extended methodology for more information on model selection, controls and post estimation tests.]. We interpret the plots and model considering multiple Amazon related events and policies over the last 30 years, as well as their correlations with different presidents and locations, while embedding problem-construction in contemporary happenings.

@@

[^2]: The reasoning here is that knowledge of high or low deforestation rates might drive presidents to speak about the Amazon. Matching deforestation with speech dates is a complicated endeavor. INPE published consolidated deforestation data with almost two years of delay until 2005 [@capobianco2021, pp. 60]. Preliminary data was published earlier, but with lower degrees of confidence in the findings. For the subsequent years, the consolidated figures for a given year tend to come out in August of the subsequent year. However, it also seems to be the case that the executive government have access to the data before everyone else. In addition, other sources indicating if deforestation is going up or down, as fire data or lower resolution deforestation data (DETER for example), circulates within the same year. The year of 1988 is particularly indicative of these complications: a report dating 1988 was circulating with deforestation and fire figures [@fearnside1990], and a New York Times article about the issue was written with comments by an INPE scientist [@simons1988]. Hence, we believe the most appropriate way of matching is without any lags on deforestation.

@@

We now move to mixed types, which average at around 21% for all presidents in our sample: overall, presidents prefer pure problem-constructions. While there is some variation in time for each single mixed type, some of them have low counts and interpretations are not adequate. We focus our discussion on those with higher incidence. First, the most frequent mix overall is that of economic integration with conservation, which after reaching 25% of all problem-constructions in 1989, remained stable at 9% on average for the remainder of the period. President Lula was the most frequent user of this mix. Second, except for economic and social development appearing together in all statements made by Color in 1992, mixed types using conservation were quite frequent in the lead up and aftermath of the 1992 Earth Summit. This includes the mix type we label sustainable development, which constructs the Amazon as a problem of economic integration, social development, and environmental conservation. In all, we interpret the appearance of mixed types as more complex understandings of Amazonian problems. This follows a global agenda of understanding interconnections of social, environmental, and economic domains. As we show that Amazonian incidence in discourse does respond to global issues, this is not a surprise given agendas as Millennium Development Goals and the Sustainable Development Goals. Nevertheless, as in pure types, we also observe the comeback of sovereignty being used in mixed types in detriment of conservation. This becomes more apparent in a comparison between Lula and Bolsonaro, the two presidents that mix the most with proportions 11% above presidential averages: 29.5% and 31.3% respectively. While the former frequently mixed conservation with other problem constructions, the latter prefers mixing sovereignty. The combination of sovereignty with economic integration, which was also characteristic of the military dictatorship policies for the region, reaches its highest level with Bolsonaro.

```{r Figure 4: mixed-types by president , echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mean_mt <- filter(comp_types, mx_cat2 == "Mixed-types") %>% mean(perc) # 21%
mixed_types_lula <- filter(comp_types, president == "Lula") # 29.5%
mixed_types_bolsonaro <- filter(comp_types, president == "Bolsonaro")# 31.3%
mixed_types_presid <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(president, mx_cat) %>%
  group_by(president) %>%
  mutate(perc = n / sum(n)) %>%
  mutate(president = factor(president,
                            levels = c("Sarney", "Collor", "Itamar", "FHC",
                                       "Lula", "Dilma", "Temer", "Bolsonaro"))) %>%
  ggplot(aes(x=president, y=perc, fill=mx_cat,label= scales::percent(perc, accuracy=.1))) +
  geom_bar(stat="identity", position="stack") +
  geom_text(size = 3, position = position_stack(vjust=.5)) +
  scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "", title ="Share of each mixed type problem-constructions by president",
       subtitle = "Sample composed of Amazonian statements that were coded as a mixed-type (n=431)",
       caption = "") +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
mixed_types_presid
```

@@

Table 2, below, displays the multinational regression coefficients for the correlation between diverse problem constructions and locations. For locations, the reference category is Amazonian states and, for problem construction, the reference category is environmental conservation. From the outset, we notice that problem constructions of pure economic integration and pure social development correlate in negative statistically significant ways to both international settings and Brasilia in relation to environmental conservation problem constructions in Amazonian states. These findings indicate that constructing the amazon as an issue of  environmental conservation is more likely to happen in international settings and in Brasilia than in Amazonian states. Additionally, pure sovereignty constructions correlate in negative statistically significant ways to international settings in relation to environmental conservation constructions in Amazonian states. That is, constructing the Amazon as an issue of sovereignty

```{r multimonial model, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
adr <- data.frame(year = c(1985, 1986, 1987), def_year = c(21.050, 21.050, 21.050))
amazon_def_year <- rbind(adr, amazon_def_year) # data on deforestation rates per year (PRODES)
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # average inflation rates per year Brazil
# remove non-identified locations, mixed-other constructions and add election years
ama_model <- filter(ama_model, location_cat != "Non Identified", mx_cat != "Other Mixed-types") %>% 
  mutate(mx_cat = factor(mx_cat, levels = c("Pure Conservation", "Pure Economic Integration",
                                            "Pure Social Development", "Pure Sovereignty",
                                            "Economic Conservation",
                                            "Economic and Social Development", "Sovereign Conservation",
                                            "Sustainable Development","Sovereign Integration",
                                            "Social Conservation", "Other")),
         election_year = ifelse(grepl("1989|1994|1998|2002|2006|2010|2014|2018", year), 1, 0))
# model
model <- multinom(mx_cat ~ location_cat + election_year + def_year + AAI, data = ama_model, model = TRUE)
# summary statistics
model_summary <- summary(model)
p_values <- round((1 - pnorm(abs(summary(model)$coefficients/summary(model)$standard.errors)))*2, 5)
# post estimation tests
chi_square <- chisq.test(ama_model$mx_cat, predict(model)) 
# Chi-square p value (p-value < 0.001) tells us that our model explains a
# significant amount of the variation in the data in comparison with a null model.
# null_model <- multinom(mx_cat ~ 1, data = ama_model) # Null model for a likelihood ration test, just in case.
# library(lmtest)
# lrtest(null_model, model) # We very much reject the hypothesis
# (i.e. the more complex model is significantly better than the null model).
# Finally, let's get a nice output:
stargazer::stargazer(model, type="text",
                     covariate.labels = c("Brasilia", "International", "Non-AM state",
                                          "Election", "Deforestation", "Inflation"),
                     column.labels = c("EI", "SD", "Sov", "EI-Con", "EI-SD", "Sov-Con", "SD-EI-Con",
                                       "Sov-EI", "SD-Con", "Other"),
                     dep.var.labels.include = FALSE)
```
